---
title: "rcrtan"
author: "Geoffrey T. LaFlair"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
    %\VignetteEngine{knitr::rmarkdown}
    %\VignetteIndexEntry{rcrtan}
    \usepackage[utf8]{inputenc}
---

This package contains functions for criterion-referenced test and item analysis. 

## Test Analysis: Dependability

The functions below can be used to examine the dependability of _single administration_ CRT assessments. 

### subkoviak

The `subkoviak` function follows Subkoviak's (1988) single administration procedures for estimating $\kappa$ and *p~o~*. To use the `subkoviak` function the data can come in one of two forms:

1. A dataframe of dichotomously scored items
2. A dataframe of total (raw) scores on the test. This also requires the analyst/researcher to know how many (dichotomous) items were on the test.

In both situations, the analyst/researcher needs to know the raw cut score of the test.

The procedure uses two pieces of information to estimate $\kappa$ and *p~o~*. They are a reliability coefficient (estimated in this function using Kuder-Richardson 20 or 21) and a standardized cut-score, which is estimated following Subkoviak (1988). The estimated $\alpha$ coefficient and standardized cut-score are used to look up the values for $\kappa$ and *p~o~* in look up tables, which the function will return if `look_up = TRUE`

$$ z = \frac{c - 0.5 - M}{S}$$

where:

- _z_ = The standardized cut-score
- _c_ = The raw cut-score of the test
- _M_ =  The mean raw total score of the test
- _S_ = The standard deviation of the raw total scores of the test

#### Examples: With item information

```{r, echo=FALSE, warning=FALSE, message=FALSE, error=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", echo=TRUE, warning=FALSE, message=FALSE, error=FALSE)
```

```{r}

# Load the library
library(rcrtan)

# Some data
test_data <- tibble::tribble(
                 ~ID, ~Q1, ~Q2, ~Q3, ~Q4, ~Q5, ~Q6, ~Q7, ~Q8, ~Q9, ~Q10, ~Total,
               1441L,  1L,  1L,  1L,  1L,  1L,  1L,  1L,  1L,  1L,   1L,    10L,
               1387L,  1L,  1L,  1L,  1L,  1L,  0L,  0L,  1L,  1L,   1L,     8L,
               1994L,  1L,  1L,  1L,  1L,  1L,  0L,  0L,  1L,  1L,   1L,     8L,
               1453L,  1L,  1L,  1L,  1L,  0L,  1L,  0L,  1L,  1L,   1L,     8L,
               1679L,  1L,  0L,  1L,  0L,  0L,  1L,  1L,  1L,  1L,   1L,     7L,
               1899L,  1L,  1L,  1L,  1L,  0L,  1L,  0L,  1L,  1L,   1L,     8L,
               1631L,  1L,  1L,  1L,  1L,  1L,  1L,  0L,  1L,  1L,   1L,     9L,
               1894L,  1L,  0L,  1L,  1L,  1L,  1L,  1L,  1L,  1L,   1L,     9L,
               1206L,  0L,  0L,  1L,  1L,  0L,  1L,  0L,  1L,  0L,   1L,     5L,
               1163L,  1L,  1L,  1L,  1L,  1L,  0L,  0L,  1L,  1L,   1L,     8L,
               1964L,  1L,  1L,  1L,  1L,  1L,  1L,  1L,  1L,  1L,   1L,    10L,
               1050L,  1L,  1L,  1L,  1L,  1L,  1L,  1L,  1L,  1L,   1L,    10L,
               1034L,  1L,  1L,  1L,  1L,  1L,  1L,  1L,  1L,  1L,   1L,    10L,
               1826L,  1L,  0L,  1L,  1L,  1L,  1L,  1L,  1L,  1L,   1L,     9L,
               1973L,  1L,  1L,  1L,  1L,  0L,  1L,  0L,  1L,  1L,   1L,     8L,
               1936L,  0L,  1L,  1L,  1L,  1L,  1L,  1L,  1L,  1L,   0L,     8L,
               1114L,  1L,  0L,  0L,  1L,  1L,  0L,  0L,  0L,  1L,   1L,     5L,
               1181L,  0L,  1L,  1L,  1L,  0L,  0L,  1L,  1L,  1L,   1L,     7L,
               1917L,  1L,  1L,  1L,  1L,  0L,  1L,  0L,  1L,  1L,   1L,     8L
               )


# Analyze the dichomous data. Use look_up = TRUE if you want to see the look up tables from Subkoviak (1988).
sub_ex_one <- subkoviak(data = test_data, items = 2:11, raw_cut_score = 8)
```

The output from the analysis contains five variables.

1. _z_: The actual estimated standardized cut-score 
2. *z_rounded*: The rounded standardized cut-score used in the table look up
3. *KR_est*: The estimated alpha coefficient using one the Kuder-Richardson methods
4. *agree_coef.r_0.4*: The estimated agreement (*p~o~*) coefficient 
5. *kappa_coef.r_0.4*: The estimated kappa ($\kappa$) coefficient

```{r}
sub_ex_one
```

#### Examples: Without item information

When item-level information is not available the _items_ argument takes the number of dichotomous items on the test, and the _total_ argument takes the name of the column in the dataframe with the total score.

```{r}
sub_ex_two <- subkoviak(data = test_data, items = 10, raw_cut_score = 8, total = "Total")

sub_ex_two
```

### phi_domain

The `phi_domain` function calculates Brown (1988) short-cut estimate of the generalizability coefficient ($\phi$ dependability) for absolute decisions.

```{r}
# When item level information is known
phi_d_one <- phi_domain(data = test_data, items = 2:11)

# When only total scores and number of items on the test are known
phi_d_two <- phi_domain(data = test_data, items = 10, total = "Total")

phi_d_one

phi_d_two
```

### phi_lambda

The `phi_lambda` function uses Brennan (1984) formula to estimate the dependability of a test. One difference to note between this function and `subkoviak` is that the *cut_score* argument is expressed as a proportion.

```{r}
phi_l_one <- rcrtan::phi_lambda(test_data, 2:11, cut_score = 0.80)

phi_l_two <- phi_lambda(test_data, 10, cut_score = 0.80, total = 'Total')

phi_l_one

phi_l_two
```

# References

Brennan, R. L. (1984). Estimating the dependability of the scores. In _A guide to criterion-referenced test construction_, (Ed.) Ronald A Berk. 231–66. Baltimore, MD: The Johns Hopkins University Press.

Brown, J. D. (1988). Short cut estimates of criterion-referenced test reliability. _Language Testing_, _7_, 77–97.

Subkoviak, M. J. (1988). A practitioner’s guide to computation and interpretation of reliability indices for mastery tests. _Journal of Educational Measurement, 25_, 47–55.